Supplementary Material

### 2025-10-09 manuscript.tex
updated manuscript files

The use of quantiles in statistical modeling and in reporting inferential or 
predictive uncertainty is widespread, and reporting several quantiles or 
predictive intervals is common for probabilistic forecasting 
\cite[]{gneiting2023model}. Quantiles can be easier to interpret than 
statistical model parameters and are often used to define confidence or 
prediction intervals. Thus with multiple quantiles for a particular outcome, 
one has a measure of uncertainty.
Estimating quantiles in the presence of covariates via 
quantile regression is a common
statistical and machine learning strategy, and where parametric models are 
complicated or nonexistent quantiles may be easier to estimate via machine 
learning methods \cite[]{martin2022direct,chung2021beyond, koenker2017quantile, 
koenker1978regression}. In estimating multiple quantiles, quantile regression 
often provides a tradeoff with parametric modeling in that for some problems 
quantile regression may be easier to perform but that predicted quantiles lack 
the detailed information of a fully defined predictive distribution 
\cite[]{pohle2020murphy}. 
% Quantile regression may also lead to other issues 
% such as quantile crossing where an estimated quantile may be smaller than 
% another estimated quantile with for a lower probability level 
% \cite[]{chernozhukov2010quantile,he1997quantile}.
% Data is often reported as quantiles, 
Perhaps, for data privacy reasons, data quantiles are often reported as 
summaries of the data.
% along with predictive distributions. 
% \jarad{Predictive distributions for what?}
Census or medical data, which can be very large or personal to the subjects, 
may be published as summary or aggregate data including percentiles (quantiles) 
and medians 
\cite[]{simpson2023interpolating,cdc2022growthcharts,nirwan2020bayesian}. In 
collaborative forecast initiatives and competitions, probabilistic forecasts 
are often submitted as a set of predictive quantiles 
% for given probabilities 
\cite[]{gneiting2023model,hong2016probabilistic}. In recent years, disease 
outbreak forecast hubs require that all forecasts submitted by outside 
participants be represented by  several predictive intervals or quantiles. 
This standardized representation allows for straightforward forecast scoring 
and ensemble building 
\cite[]{mathis2024evaluation,mathis2023flusight,Cramer2022-hub-dataset,
cramer2022evaluation,sherratt2023predictive,bracher2021evaluating}.

@@

In this manuscript we introduce a novel model, the quantile Gaussian process 
(QGP), used specifically for QM. The context for using the QGP is that the 
available data is a set of quantiles estimated for a given set of 
probabilities.
The QGP relies on established theory underlying the relationship between the 
QF of a continuous distribution and sample quantiles 
\cite[]{parzen2004quantile, gilchrist2000statistical,hyndman1996sample,
walker1968note,cramer1951mathematical}. 
The QGP can provide accurate asymptotic inference of the quantile uncertainty, 
and it works well for QM when modeling quantile defined distributions. 
In section \ref{sec:quant_func} the QF of a continuous random variable and 
sample quantiles are each defined, and important properties and asymptotic 
theory are reviewed. In section \ref{sec:qgp_model} the QGP model for QM is 
introduced. Section \ref{sec:simulation_analyses} contains simulation studies 
illustrating the QGP's ability to make parameter inference and match the 
distribution from which the quantiles where estimated. Section 
\ref{sec:qm_analysis} contains an application of QM using the QGP model for 
quantile forecasts of the 2023-24 United States Centers for Disease Control and
Prevention
(CDC) flu forecasting competition, also known as FluSight. Section 
\ref{sec:qm_conclusion} concludes this manuscript with a short summary and 
ideas for further development in QM modeling.

@@

Along with the CDF and PDF, the QF is a defining function of a random variable. 
The QF, however, often receives less attention in standard statistical 
training than do the CDF and PDF \cite[]{parzen2004quantile}, and the CDF and 
PDF are used more often for modeling and inference.
% But the QF tends to be the little brother 
% \jarad{little brother?}
% of both the CDF and PDF in that the latter two functions are taught and used far more in statistical modeling and inference. 
% Nevertheless, the QF and several of its properties are central to the QGP model introduced in this paper. 
% \jarad{another throwaway sentence}
This section contains the definition of the QF, important properties, the 
definition of sample quantiles, and key central limit theorems (CLTs) of sample 
quantiles.

@@

A number of metrics for measuring the distance between two continuous 
univariate distributions exist. Those we consider in this paper are the 
Wasserstein distance (WD) and a slight modification of it, the total variation 
(TV) or the statistical difference, and the Kullback-Leibler divergence (KLD). 
A brief overview of these and other metrics, some of their statistical uses 
and properties, and relationships between metrics is found in 
\cite{gibbs2002choosing}. These metrics were used in section 
\ref{sec:simulation_analyses} to assess how well different QM methods 
approximate a true distribution.

@@

\subsection{Parameter estimation and quantile matching for known distribution \\families}

@@

\subsubsection{Normal parameter estimation}
In this simulation study, we fit the ORD and IND models to quantiles estimated 
from independent samples from a normal distribution and compare parameter 
estimation with fits from the QGP model. Quantiles were simulated by first 
drawing a sample of size $n$ from a known distribution then estimating $K$ 
quantiles given probabilities $\{p_1, ..., p_K\}$. The ORD, IND, and QGP models 
were then fit to the quantiles where $F$ was set to be the CDF of a normal 
distribution. For each of 
$n \in \{50; 150; 500; 1{,}000; 5{,}000\}$ and 
% $K \in \{3,7,15,50\}$
$K = 23$, there 
were 500 simulation replicates. The data were simulated from a normal 
distribution with mean $\mu = 4$ and standard deviation $\sigma = 3.5$.
For each of QGP, ORD, and IND models, we assigned the same prior distributions 
to $\mu$ and $\sigma$. When $n$ is unknown, we also assigned it a prior 
distribution. 
As one will generally have access to the quantiles being fit, 
it is reasonable that a informative prior $\mu$ centered near where the data is
centered be assigned. We took this approach herein, and for spread parameters,
we assigned normal priors truncated at 0 as recommended by
\cite{gelman2006prior}.
The prior distributions were $\mu \sim N(5,7^2)$,
$\sigma \sim N(0, 6^2) \mathds{1} \{\sigma > 0\}$, and
$n \sim N(0,3000^2)\mathds{1}\{n > 0\}$.
% \begin{align*}
%     \mu &\sim N(5,7^2) \\
%     \sigma &\sim N(0, 6^2) \mathds{1}\{\sigma > 0\}\\
%     n &\sim N(0,3000^2)\mathds{1}\{n > 0\}.
% \end{align*}
For the IND model the prior on the independent parameter 
$\sigma_{\rho}$ from model (\ref{eq:ind_quantile}) is 
$1/\sigma_{\rho} \sim N(0, 3000^2)\mathds{1}\{\sigma_{\rho} > 0\}$,
which is similar to the prior on $n$ for 
the QGP and ORD models. For each fit, the HMC chain was run for 60,000 draws 
with the first 10,000 discarded as a burn-in.

@@

The top of figure \ref{fig:normal_cov_dists} shows the 90\% coverage of 
the posterior distribution credible intervals as $n$ increases for 
$\mu$, $\sigma$ and $n$ for five models. QGP-n and ORD-n models where $n$ is 
known are included along with QGP, ORD, and IND. 
The 90\% credible intervals of the 
parameters were calculated by computing the $0.05^{th}$ and $0.95^{th}$ 
quantiles from the posterior distribution samples. The coverage percentage is 
calculated as the percentage of the 500 replicates which the true parameter 
value was within the 90\% credible interval.
As expected, the nominal coverage for the QGP and ORD models is better than 
that of the IND model, and for the two models where $n$ is known the coverage 
is better than where it is unknown, particularly for lower values of 
$n$ and $K$. The coverage for the ORD models appears to be slightly better 
than for the QGP models though not by much, and the difference
decreases as $n$ increases. 
The superior performance of the ORD model is possibly because it
provides exact inference 
whereas the QGP model provides asymptotic inference.

@@

With the true distribution family being unknown, we made the modeling decision 
to set $F$ in the QGP model of (\ref{eq:pit_qgp}) to be a mixture of normals 
distribution taking the form of (\ref{eq:norm_mix}). This decision to use
a normal mixture was 
based on the common claim that any continuous distribution may be approximated 
by a finite mixture of normal distributions 
\cite[]{peel2000finite,nguyen2019approximations,nguyen2020approximation}. This 
claim may be optimistic, but we found that a four component normal mixture 
distribution can approximate the non-normally shaped distributions reasonably 
well. However, any function which meets the technical definition of a CDF would 
be appropriate to use for $F$, see \cite{gasthaus2019probabilistic} for an 
example.
After a small analysis comparing fits of one to six normal components,
the number of component 
distributions was chosen to be $C = 4$. The results, included in the 
supplementary materials, show that as the number of components in the mixture
distrubtion increases, the distribution approximation improves. But for the 
cases considered, the improvement from four to five components or more is small.

The three distributions from which data were simulated and sample quantiles 
estimated and to which QM methods were fit were the extreme value (EV) 
distribution with location 0 and scale 1 $EV(0,1)$, Laplace (La) with location 
0 and scale 1 $La(0,1)$, and a two component normal mixture (MIX) 
$w N(-1, 0.9) + (1-w)N(1.2, 0.6)$ where $w = 0.35$. From the three 
distributions, 500 replicates of $K = 23$ quantiles were simulated for each of 
$n \in \{50, 150, 500, 1{,}000, 2{,}000, 5{,}000\}$. QM for each replicate of 
simulated quantiles was done using QGP, ORD, IND, SPL, and KDE. For the QGP, 
ORD, and IND models, $\theta = \{\theta_1, \theta_2, \theta_3, \theta_4\}$ 
where $\theta_c = (\mu_c, \sigma_c)$ for $c \in \{1,2,3,4\}$. The prior 
distributions assigned were $\mu_c \overset{ind}{\sim} N(5, 7^2)$,
$\sigma_c \overset{ind}{\sim} N(0, 6^2)\mathds{1}\{\sigma_c > 0\}$,
$w \sim Dirichlet(\boldsymbol{1}_C)$, and
$\nu \sim N(0, 3000^2)\mathds{1}\{\nu > 0\}$
% as below 
where $\nu = n$ for the QGP and ORD models 
and $\nu = 1/\sigma_{\rho}$ for the IND model. 
% (cite an example of using dirichlet prior ??). 
    % \begin{align*}
    %     \mu_c &\overset{ind}{\sim} N(5, 7^2) \\
    %     \sigma_c &\overset{ind}{\sim} N(0, 6^2)\mathds{1}\{\sigma_c > 0\} \\ 
    %     w &\sim Dir(\boldsymbol{\alpha}), \quad \alpha_c = 1 \\
    %     \nu &\sim N(0, 3000^2)\mathds{1}\{\nu > 0\}
    % \end{align*}
 %the problem with $\nu$ is that for the QGP and ORD the prior is on the variance whereas for IND it is on the sd.
Often when implementing posterior sampling for a mixture distribution model, 
one deals with an issue called the label-switching problem. This is where for a
mixture distribution parameter $\theta = \{\theta_1,...,\theta_C\}$, the model 
likelihood is the same for different permutations of $\theta$. This lack of 
identifiability for elements of $\theta$ makes parameter inference useless, 
but the predictive distribution may still be close to the true distribution 
\cite[]{stephens2000dealing}.
Because of this, the HMC posterior sampling chain was run longer than in the 
studies in the previous section. The chain was run for 80,000 steps where the 
first 20,000 draws were discarded as a burn-in. For this study, we were 
concerned only with QM and not with parameter inference, so the fact that the 
parameters are unidentifiable was not considered critical. However, when 
assessing the TV and KLD between the QM and true distributions, variational 
Bayes (VB) in \texttt{Stan} was used instead of HMC for fitting the models 
\cite[]{kucukelbir2015automatic}. This decision was made because of the lack of 
parameter identifiability and because the marginal parameter means are used to 
estimate the PDF. While MCMC methods perform better on parameter inference, 
VB methods have been shown to have predictive performance comparable to MCMC 
\cite[]{blei2017variational}.

Figures \ref{fig:evd_fits}, 
shows examples of fits of $K=23$ quantiles simulated from the EV distribution
for different sample sizes $n$. Included QM methods 
are KDE, SPL, IND, and QGP. ORD was excluded to make visualization easier, but 
the fits are very similar to the QGP fits. The KDE and SPL fits provide no 
uncertainty estimation of the quantiles, and return only a continuous function. 
The SPL fits show a lot of wiggle with the wiggle decreasing as $n$ increases, 
whereas the KDE fits do not show as much wiggle, but the KDE fits the 
distribution tails poorly even as $n$ increases. The IND and QGP models provide 
uncertainty in the fits, but the QGP is much more conservative with wider 
intervals which provide superior coverage of the quantiles and the PDFs.

@@

The top of figure \ref{fig:uk_cover} shows the 95\% simulated coverage of the 
true quantiles 
for QGP, ORD, and IND fits of quantiles from the EV, La, and MIX distributions. 
The percent coverage is 
averaged over all $K = 23$ quantiles and all 500 simulated replicates.  
The plots show that the QGP and ORD models are largely able to meet and exceed 
the nominal coverage where the IND model falls very short. Because it took a 
long time to evaluate the QF when calculating coverage, calculation was done on 
a thinned sample of the posterior distribution where only every $100^{th}$ HMC 
draw was kept so that the coverage was calculated using 6,000 posterior draws 
rather than the full 60,000.

@@

The bottom of \ref{fig:uk_cover} shows the UWD1, TV, and KLD distances between 
QM and true distributions averaged over 500 simulation replicates for the 
five QM methods. In the UWD1 plot, there appears to be little or no difference
in performance between the QGP, ORD, IND, and SPL fits, but each of these
outperforms the KDE fits. For TV, the QGP, ORD, and SPL perform similarly, 
though the QGP does the best for the MIX distribution.
And for KLD, the QGP and ORD perform the best for lower values of $n$, but as
$n$ increases, SPL appears to perform the best. Additional plots and results
from these analyses are found in the supplementary material.

@@

This section shows the ability of the QGP model to accurately perform parameter 
inference and QM when given sample quantiles. Although the inference for QGP 
is asymptotic, the QGP performance in inference and in QM is similar to that 
of ORD where the inference is exact. There are also circumstances where the 
QGP may outperform the ORD as shown in cases where the QF is easily evaluated 
but where the CDF is difficult to evaluate. Another case where the QGP 
outperformed the ORD was in QM of the normal mixture distribution. An 
advantage that the QGP has over SPL and KDE is that it provides accurate 
uncertainty estimation of quantile uncertainty where SPL and KDE provide no 
uncertainty estimation.

@@

Participating teams were free to create forecasts however they pleased, but 
forecasts for each target were required to consist of 23 quantiles for the 
probability levels 
$\boldsymbol{p} =(0.01, 0.025, 0.5, 0.1, ..., 0.95, 0.975, 0.99)$, 
which were given by FluSight. Figure \ref{fig:quant_forcs_us} shows examples 
of quantile forecasts of the same target from 12 participating teams. Each plot 
shows the log forecast for hospitalizations in the US for the week of 
January 13, 2024, and it is clear that there are major distributional 
differences between forecasts.
We denote a quantile forecast for flu hospitalizations as 
$\hat{Q}^{(H)}(\boldsymbol{p}) = (\hat{Q}^{(H)}(0.01), ..., \hat{Q}^{(H)}(0.99))$. 
The general quantile forecast representation format allowed for forecasts to be 
compared by the same metric, the weighted interval score (WIS), and to be 
easily combined into a multi-model ensemble forecast 
\cite[]{mathis2024evaluation}. However, under the quantile representation, 
the tools for scoring forecasts and for building ensemble models are limited 
as many of the existing tools for scoring forecasts and constructing ensembles 
require CDFs or PDFs \cite[]{wadsworth2023mixture,ranjan2010combining}. 
Because of these limitations, it may be desirable to approximate continuous 
distributions from the quantile forecasts to allow for more flexibility in 
scoring or ensemble building. In this section, we fit the QGP and other
models for QM to 
forcasts submitted to the FluSight during the 2023-24 
season, and an analysis is made to compare the fits.

@@

Before fitting a QGP model to a quantile forecast we transformed the quantiles 
$\hat{Q}^{(H)}(\boldsymbol{p})$ to $\text{log}(\hat{Q}^{(H)}(\boldsymbol{p}) + 1)$ 
so that forecasts for all states were on a similar scale. The model fit was 
the same as the QGP fit in section \ref{sec:uk_matching}, including the same 
prior distribution assignments.
According to the official forecast competition rules, no forecast could include 
quantiles that were less than 0. As a result, there were forecasts with one or 
many quantiles equal to 0. When fitting the QGP model, these instances of 0 
values were removed so that some forecasts had $K < 23$ quantiles. Between 39 
competing forecast teams, 53 locations, 29 forecast dates, and 4 horizons, 
there were 161,093 quantile forecasts on which QM was done.
For each 
forecast, the WIS was calculated for the quantile forecast and the continuous 
ranked probability score (CRPS) was calculated over the predictive distribution 
of the fit QGP model.

@@

Figures \ref{fig:flu_methods_corr} shows results of 
comparing the WIS from quantile forecasts and the CRPS from the 
fitted forecasts for QGP, ORD, IND, SPL, and KDE methods. 
The plots have in the $x$-axis the WIS and 
CRPS is in the $y$-axis. Each point is for one forecast for any participating 
team, season week, and state. Unsurprisingly, the correlation between the
QM CRPS and WIS is very highly correlated for each method. For all but the KDE
method, there is a tendency for some CRPS values to go above the $x = y$ line
suggesting the CRPS scores are enlarged due to the addition of distribution
tails. The CRPS scores for the IND method also go below the $x = y$ line 
suggesting that the method's fit may be simply be poorer than the other methods.
The slightly lower correlation for the QGP compared to ORD and IND may also be
a result of adding more uncertainty in the fitting of tail quantiles.

@@

%% If you have bib database file and want bibtex to generate the
%% bibitems, please use
%%
% \bibliographystyle{sn-apacite}
\bibliography{master_bib}

>>>>

The use of quantiles in statistical modeling and in reporting inferential or 
predictive uncertainty is widespread, and reporting several quantiles or 
predictive intervals is common for probabilistic forecasting 
\cite[]{gneiting2023model}. Quantiles can be easier to interpret than 
statistical model parameters and are often used to define confidence or 
prediction intervals. Thus with multiple quantiles for a particular outcome, 
one has a measure of uncertainty.
Estimating quantiles in the presence of covariates via 
quantile regression is a common
statistical and machine learning strategy, and where parametric models are 
complicated or nonexistent quantiles may be easier to estimate via machine 
learning methods \cite[]{martin2022direct,chung2021beyond, koenker2017quantile, 
koenker1978regression}. In estimating multiple quantiles, quantile regression 
often provides a tradeoff with parametric modeling in that for some problems 
quantile regression may be easier to perform but that predicted quantiles lack 
the detailed information of a fully defined predictive distribution 
\cite[]{pohle2020murphy}. 
% Quantile regression may also lead to other issues 
% such as quantile crossing where an estimated quantile may be smaller than 
% another estimated quantile with for a lower probability level 
% \cite[]{chernozhukov2010quantile,he1997quantile}.
% Data is often reported as quantiles, 
Perhaps for data privacy reasons, data quantiles are often reported as 
summaries of the data.
% along with predictive distributions. 
% \jarad{Predictive distributions for what?}
Census or medical data, which can be very large or personal to the subjects, 
may be published as summary or aggregate data including percentiles (quantiles) 
and medians 
\cite[]{simpson2023interpolating,cdc2022growthcharts,nirwan2020bayesian}. In 
collaborative forecast initiatives and competitions, probabilistic forecasts 
are often submitted as a set of predictive quantiles 
% for given probabilities 
\cite[]{gneiting2023model,hong2016probabilistic}. In recent years, disease 
outbreak forecast hubs require that all forecasts submitted by outside 
participants be represented by  several predictive intervals or quantiles. 
This standardized representation allows for straightforward forecast scoring 
and ensemble building 
\cite[]{mathis2024evaluation,mathis2023flusight,Cramer2022-hub-dataset,
cramer2022evaluation,sherratt2023predictive,bracher2021evaluating}.

@@

In this manuscript we introduce a novel model, the quantile Gaussian process 
(QGP) which is used specifically for QM.
The context for using the QGP is that the 
available data is a set of quantiles estimated for a given set of 
probabilities.
The QGP relies on established asymptotic 
theory underlying the relationship between the 
QF of a continuous distribution and sample quantiles 
\cite[]{parzen2004quantile, gilchrist2000statistical,hyndman1996sample,
walker1968note,cramer1951mathematical}. 
The QGP can provide accurate asymptotic inference of the quantile uncertainty, 
and it works well for QM when modeling quantile defined distributions. The 
QGP is fit using Bayesian methods and thus allows for using prior information
to influence the QM.
In section \ref{sec:quant_func} the QF of a continuous random variable and 
sample quantiles are each defined, and important properties and asymptotic 
theory are reviewed. In section \ref{sec:qgp_model} the QGP model for QM is 
introduced. Section \ref{sec:simulation_analyses} contains simulation studies 
illustrating the QGP's ability to make parameter inference and match the 
distribution from which the quantiles where estimated. Section 
\ref{sec:qm_analysis} contains an application of QM using the QGP model for 
quantile forecasts of the 2023-24 United States Centers for Disease Control and
Prevention
(CDC) flu forecasting competition, also known as FluSight. 
The analyses in sections \ref{sec:simulation_analyses} and
\ref{sec:qm_analysis} show that the QGP accurately estimates distribuions from
which the quantiles are estimated and while doing so accurately captures the 
uncertainty inherent in estimating quantiles. The QGP outperforms all
but one other competing methods in estimating the unknown distributions and
capturing uncertainty, and also allows for QM in situations where the other 
methods cannot be used.
Section 
\ref{sec:qm_conclusion} concludes this manuscript with a short summary and 
ideas for further development in QM modeling.

@@

Along with the CDF and PDF, the QF is a defining function of a random variable. 
The QF, however, often receives less attention in standard statistical 
training than do the CDF and PDF \cite[]{parzen2004quantile}, and the CDF and 
PDF are used more often for modeling and inference.
% But the QF tends to be the little brother 
% \jarad{little brother?}
% of both the CDF and PDF in that the latter two functions are taught and used far more in statistical modeling and inference. 
% Nevertheless, the QF and several of its properties are central to the QGP model introduced in this paper. 
% \jarad{another throwaway sentence}
This section contains the definition of the QF, important properties, the 
definition of sample quantiles, and key central limit theorems (CLTs) of sample 
quantiles on which the QGP is based.

@@

A number of metrics for measuring the distance between two continuous 
univariate distributions exist. Those we consider in this manuscript are the 
Wasserstein distance (WD) and a slight modification of it, the total variation 
(TV) or the statistical difference, and the Kullback-Leibler divergence (KLD). 
A brief overview of these and other metrics, some of their statistical uses 
and properties, and relationships between metrics is found in 
\cite{gibbs2002choosing}. These metrics were used in section 
\ref{sec:simulation_analyses} to assess how well different QM methods 
approximate a true distribution.

@@

\subsection{Parameter estimation and quantile matching for known distribution families}

@@

\subsubsection{Normal parameter estimation}
In this simulation study, we fit the ORD and IND models to quantiles estimated 
from independent samples from a normal distribution and compare parameter 
estimation with fits from the QGP model. Quantiles were simulated by first 
drawing a sample of size $n$ from a known distribution then estimating $K$ 
quantiles given probabilities $\{p_1, ..., p_K\}$. The ORD, IND, and QGP models 
were then fit to the quantiles where $F$ was set to be the CDF of a normal 
distribution. For each of 
$n \in \{50, 150, 500, 1{,}000, 5{,}000\}$ and 
% $K \in \{3,7,15,50\}$
$K = 23$, there 
were 500 simulation replicates. The data were simulated from a normal 
distribution with mean $\mu = 4$ and standard deviation $\sigma = 3.5$.
For each of QGP, ORD, and IND models, we assigned the same prior distributions 
to $\mu$ and $\sigma$. When $n$ is unknown, we also assigned it a prior 
distribution. 
As one will generally have access to the quantiles being fit, 
it is reasonable that an
informative prior, centered near where the data is
centered, be assigned to $\mu$. 
We took this approach herein, and for scale parameters,
we assigned normal priors truncated at 0 as recommended by
\cite{gelman2006prior}.
The prior distributions were $\mu \sim N(5,7^2)$,
$\sigma \sim N(0, 6^2) \mathds{1} \{\sigma > 0\}$, and
$n \sim N(0,3000^2)\mathds{1}\{n > 0\}$.
% \begin{align*}
%     \mu &\sim N(5,7^2) \\
%     \sigma &\sim N(0, 6^2) \mathds{1}\{\sigma > 0\}\\
%     n &\sim N(0,3000^2)\mathds{1}\{n > 0\}.
% \end{align*}
For the IND model the prior on the independent parameter 
$\sigma_{\rho}$ from model (\ref{eq:ind_quantile}) is 
$1/\sigma_{\rho} \sim N(0, 3000^2)\mathds{1}\{\sigma_{\rho} > 0\}$,
which is similar to the prior on $n$ for 
the QGP and ORD models. For each fit, the HMC chain was run for 60,000 draws 
with the first 10,000 discarded as a burn-in.

@@

The top of figure \ref{fig:normal_cov_dists} shows the 90\% coverage of 
the posterior distribution credible intervals as $n$ increases for 
$\mu$, $\sigma$ and $n$ for five models. QGP-n and ORD-n models where $n$ is 
known are included along with QGP, ORD, and IND. 
The 90\% credible intervals of the 
parameters were calculated by computing the $0.05^{th}$ and $0.95^{th}$ 
quantiles from the posterior distribution samples. The coverage percentage is 
calculated as the percentage of the 500 replicates for 
which the true parameter 
value was within the 90\% credible interval.
As expected, the nominal coverage for the QGP and ORD models is better than 
that of the IND model, and for the two models where $n$ is known the coverage 
is better than where it is unknown, particularly for lower values of 
$n$ and $K$. The coverage for the ORD models appears to be slightly better 
than for the QGP models though not by much, and the difference
decreases as $n$ increases. 
The superior performance of the ORD model is possibly because it
provides exact inference 
whereas the QGP model provides asymptotic inference.

@@

With the true distribution family being unknown, we made the modeling decision 
to set $F$ in the QGP model of (\ref{eq:pit_qgp}) to be a mixture of normals 
distribution taking the form of (\ref{eq:norm_mix}). This decision to use
a normal mixture was 
based on the common claim that any continuous distribution may be approximated 
by a finite mixture of normal distributions 
\cite[]{peel2000finite,nguyen2019approximations,nguyen2020approximation}. This 
claim may be optimistic, but we found that with enough components,
the normal mixture 
distribution can approximate the non-normally shaped distributions reasonably 
well. However, any function which meets the technical definition of a CDF would 
be appropriate to use for $F$, see \cite{gasthaus2019probabilistic} for an 
example.
After a small analysis comparing fits of one to six normal components,
the number of component 
distributions was chosen to be $C = 4$. The results, included in the 
supplementary materials, show that as the number of components in the mixture
distrubtion increases, the distribution approximation improves. But for the 
cases considered, the improvement from four to five components or more is small.

The three distributions from which data were simulated and sample quantiles 
estimated and to which QM methods were fit were the extreme value (EV) 
distribution with location 0 and scale 1 $EV(0,1)$, Laplace (La) with location 
0 and scale 1 $La(0,1)$, and a two component normal mixture (MIX) 
$w N(-1, 0.9) + (1-w)N(1.2, 0.6)$ where $w = 0.35$. From the three 
distributions, 500 replicates of $K = 23$ quantiles were simulated for each of 
$n \in \{50, 150, 500, 1{,}000, 2{,}000, 5{,}000\}$. QM for each replicate of 
simulated quantiles was done using QGP, ORD, IND, SPL, and KDE. For the QGP, 
ORD, and IND models, $\theta = \{\theta_1, \theta_2, \theta_3, \theta_4\}$ 
where $\theta_c = (\mu_c, \sigma_c)$ for $c \in \{1,2,3,4\}$. The prior 
distributions assigned were $\mu_c \overset{ind}{\sim} N(5, 7^2)$,
$\sigma_c \overset{ind}{\sim} N(0, 6^2)\mathds{1}\{\sigma_c > 0\}$,
$w \sim Dirichlet(\boldsymbol{1}_C)$, and
$\nu \sim N(0, 3000^2)\mathds{1}\{\nu > 0\}$
% as below 
where $\nu = n$ for the QGP and ORD models 
and $\nu = 1/\sigma_{\rho}$ for the IND model. 
% (cite an example of using dirichlet prior ??). 
    % \begin{align*}
    %     \mu_c &\overset{ind}{\sim} N(5, 7^2) \\
    %     \sigma_c &\overset{ind}{\sim} N(0, 6^2)\mathds{1}\{\sigma_c > 0\} \\ 
    %     w &\sim Dir(\boldsymbol{\alpha}), \quad \alpha_c = 1 \\
    %     \nu &\sim N(0, 3000^2)\mathds{1}\{\nu > 0\}
    % \end{align*}
 %the problem with $\nu$ is that for the QGP and ORD the prior is on the variance whereas for IND it is on the sd.
Often when implementing posterior sampling for a mixture distribution model, 
one deals with an issue called the label-switching problem. This is where for a
mixture distribution parameter $\theta = \{\theta_1,...,\theta_C\}$, the model 
likelihood is the same for different permutations of $\theta$. This lack of 
identifiability for elements of $\theta$ makes parameter inference useless, 
but the predictive distribution may still be close to the true distribution 
\cite[]{stephens2000dealing}.
Because of this, the HMC posterior sampling chain was run longer than in 
studies of the previous section. The chain was run for 80,000 steps where the 
first 20,000 draws were discarded as a burn-in. For this study, we were 
concerned only with QM and not with parameter inference, so the fact that the 
parameters are unidentifiable was not considered critical. However, when 
assessing the TV and KLD between the QM and true distributions, variational 
Bayes (VB) in \texttt{Stan} was used instead of HMC for fitting the models 
\cite[]{kucukelbir2015automatic}. This decision was made because of the lack of 
parameter identifiability and because the marginal parameter means are used to 
estimate the PDF. While MCMC methods perform better on parameter inference, 
VB methods have been shown to have predictive performance comparable to MCMC 
\cite[]{blei2017variational}.

Figure \ref{fig:evd_fits}, 
shows examples of fits of $K=23$ quantiles simulated from the EV distribution
for different sample sizes $n$. Included QM methods 
are KDE, SPL, IND, and QGP. ORD was excluded to make visualization easier, but 
the fits are very similar to the QGP fits. The KDE and SPL fits provide no 
uncertainty estimation of the quantiles, and return only a continuous function. 
The SPL fits show a lot of wiggle with the wiggle decreasing as $n$ increases, 
whereas the KDE fits do not show as much wiggle, but the KDE fits the 
distribution tails poorly even as $n$ increases. The IND and QGP models provide 
uncertainty in the fits, but the QGP is much more conservative with wider 
intervals which provide superior coverage of the quantiles and the PDFs.

@@

The top of figure \ref{fig:uk_cover} shows the 95\% simulated coverage of the 
true quantiles 
for QGP, ORD, and IND fits of quantiles from the EV, La, and MIX distributions. 
The percent coverage is 
averaged over all $K = 23$ quantiles and all 500 simulated replicates.  
The plots show that the QGP and ORD models are largely able to meet and exceed 
the nominal coverage where the IND model coverage is low. Because of the time it
took to evaluate the QF when calculating coverage, calculation was done on 
a thinned sample of the posterior distribution where only every $100^{th}$ HMC 
draw was kept so that the coverage was calculated using 6,000 posterior draws 
rather than the full 60,000.

@@

The bottom of \ref{fig:uk_cover} shows the UWD1, TV, and KLD distances between 
QM and true distributions averaged over 500 simulation replicates for the 
five QM methods. In the UWD1 plot, there appears to be little or no difference
in performance between the QGP, ORD, IND, and SPL fits, but each of these
outperforms the KDE fits. For TV, the QGP, ORD, and SPL perform similarly, 
though the QGP does the best for the MIX distribution.
And for KLD, the QGP and ORD perform the best for lower values of $n$, but as
$n$ increases, SPL appears to perform the best. Additional plots and results
from these analyses are found in the supplementary materials.

@@

This section shows the ability of the QGP model to accurately perform parameter 
inference and QM when given sample quantiles. Although the inference for QGP 
is asymptotic, the QGP performance in inference and in QM is similar to that 
of ORD which provides exact inference. There are also circumstances where the 
QGP may outperform the ORD as shown in cases where the QF is easily evaluated 
but where the CDF is difficult to evaluate. Another case where the QGP 
outperformed the ORD was in QM of the normal mixture distribution. An 
advantage that the QGP has over SPL and KDE is that it provides accurate 
uncertainty estimation of quantile uncertainty where SPL and KDE provide no 
uncertainty estimation.

@@

Participating teams were free to create forecasts however they pleased, but 
forecasts for each target were required to consist of 23 quantiles for 
probability levels 
$\boldsymbol{p} =(0.01, 0.025, 0.5, 0.1, ..., 0.95, 0.975, 0.99)$, 
which were given by FluSight. Figure \ref{fig:quant_forcs_us} shows examples 
of quantile forecasts of the same target from 12 participating teams. Each plot 
shows the log forecast for hospitalizations in the US for the week of 
January 13, 2024, and it is clear that there are major distributional 
differences between forecasts.
We denote a quantile forecast for flu hospitalizations as 
$\hat{Q}^{(H)}(\boldsymbol{p}) = (\hat{Q}^{(H)}(0.01), ..., \hat{Q}^{(H)}(0.99))$. 
The general quantile forecast representation format allowed for forecasts to be 
compared by the same metric, the weighted interval score (WIS), and to be 
easily combined into a multi-model ensemble forecast 
\cite[]{mathis2024evaluation}. However, under the quantile representation, 
the tools for scoring forecasts and for building ensemble models are limited 
as many of the existing tools for scoring forecasts and constructing ensembles 
require CDFs or PDFs \cite[]{wadsworth2023mixture,ranjan2010combining}. 
Because of these limitations, it may be desirable to approximate continuous 
distributions from the quantile forecasts to allow for more flexibility in 
scoring or ensemble building. In this section, we fit the QGP and other
models for QM to 
forcasts submitted to the FluSight during the 2023-24 
season, and an analysis is made to compare the fits.

@@

Before fitting a QGP model to a quantile forecast we transformed the quantiles 
$\hat{Q}^{(H)}(\boldsymbol{p})$ to $\text{log}(\hat{Q}^{(H)}(\boldsymbol{p}) + 1)$ 
so that forecasts for all states were on a similar scale. The model fit was 
the same as the QGP fit in section \ref{sec:uk_matching}, including the same 
prior distribution assignments.
According to the official forecast competition rules, no forecast could include 
quantiles that were less than 0. As a result, there were forecasts with one or 
many quantiles equal to 0. When fitting the QGP model, the 0 
values were removed so that some forecasts had $K < 23$ quantiles. Between 39 
competing forecast teams, 53 locations, 29 forecast dates, and 4 horizons, 
there were 161,093 quantile forecasts on which QM was done.
For each 
forecast, the WIS was calculated for the quantile forecast and the continuous 
ranked probability score (CRPS) was calculated over the predictive distribution 
of the fit QM models.

@@

Figure \ref{fig:flu_methods_corr} shows results of 
comparing the WIS from quantile forecasts and the CRPS from the 
fitted forecasts for QGP, ORD, IND, SPL, and KDE methods. 
The plots have in the $x$-axis the WIS and 
CRPS is in the $y$-axis. Each point represents
one forecast for any participating 
team, season week, and state. Unsurprisingly, the correlation between the
QM CRPS and WIS is very highly correlated for each method. For all but the KDE
method, there is a tendency for some CRPS values to go above the $x = y$ line
suggesting the CRPS scores are enlarged due to the addition of distribution
tails. The CRPS scores for the IND method also go below the $x = y$ line 
suggesting that the method's fit may be simply be poorer than the other methods.
The slightly lower correlation for the QGP compared to ORD and IND may also be
a result of adding more uncertainty in the fitting of tail quantiles.

@@

%% If you have bib database file and want bibtex to generate the
%% bibitems, please use
%%
% \bibliographystyle{sn-mathphys}
\bibliography{master_bib}
